---
title: "Model details"
author: "Graeme Baird"
date: "5/24/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../codefiles/Functions.R")
library(cowplot)
library(bayesplot)
library(magrittr)
```

### Underlying model details

In this paper we present an extension of a cumulative link ordinal regression model to include nonlinear effects, and it's application to determining chill sufficiency in cherry production. Below are figures illustrating the effect of variations in parameters.

### Plotting response of nonlinear CP equation to changes in parameters (X:CP, Y:Coefficient output)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
out <- 100

CP <- seq(0,70,length.out=out)


logfun <- function(CP, alpha, beta, eta) {
  lout <- alpha/(1+exp(-beta - CP*eta))
  return(lout)
}

logmak <- function(CP, alpha, beta, eta){
  funout <- rep(NA,out)
  for(i in 1:out){funout[i] <- logfun(CP[i], alpha, beta, eta)}
  return(funout)
}

logit_pred <- function(GDH, CP, alpha, beta, eta) {
  ak <- c(2,5)
  phi <- GDH * logfun(CP, alpha, beta, eta);
  storeframe <- pordlogit(1:3, a=ak, phi=phi)
  storeframe[3] <- storeframe[3]- storeframe[2]
  storeframe[2] <- storeframe[2] - storeframe[1]
  return(1-storeframe[1])
}

logit_mak <- function(GDH, CP, alpha, beta, eta){
  funout <- rep(NA,out)
  for(i in 1:out){funout[i] <- logit_pred(GDH[i], CP, alpha, beta, eta)}
  return(funout)
}

logit_pred_gamma <- function(GDH, CP, alpha, beta, eta) {
  ak <- c(2,5)
  GDH * logfun(CP, alpha, beta, eta);
}

logit_mak_gamma <- function(GDH, CP, alpha, beta, eta){
  funout <- rep(NA,out)
  for(i in 1:out){funout[i] <- logit_pred_gamma(GDH[i], CP, alpha, beta, eta)}
  return(funout)
}

```

Parameter variance within the non-linear component is best visualized by manipulation of individual parameters, as follows. Parameter ranges demonstrated below are appropriate for un-scaled variables and don't necessarily match actual model parameter estimates, which are fit on scaled observations. 


#### Changes in alpha determine asymptotic maximum possible growth rate 

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 5}

CP <- seq(0,70,length.out=100)
GDH <- seq(0,20000,length.out=out)

plot(CP, logfun(CP, 3,-30,1),type="l", xlim=c(0,80), ylab="")
text(75,3,labels="alpha = 3")
lines(CP, logfun(CP, 2,-30,1),type="l", col="red")
text(75,2,labels="alpha = 2")
lines(CP, logfun(CP, 1,-30,1),type="l", col="blue")
text(75,1,labels="alpha = 1")
lines(CP, logfun(CP, .5,-30,1),type="l", col="purple")
text(75,.5,labels="alpha = .5")

```

#### Changes in beta determine overall location of curve
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 5}

plot(CP, logfun(CP, 3,-40,1),type="l", xlim=c(0,70), ylab="")
text(45,2.5,labels="beta = -40")
lines(CP, logfun(CP, 3,-30,1),type="l", col="red")
text(35,2,labels="beta = -30")
lines(CP, logfun(CP, 3,-20,1),type="l", col="blue")
text(25,1.5,labels="beta = -20")
lines(CP, logfun(CP, 3,-10,1),type="l", col="purple")
text(15,1,labels="beta = -10")

```

#### Changes in eta determine location and acceleration between inflection points
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 5}

plot(CP, logfun(CP, 1,-10,.5),type="l", xlim=c(0,70), ylab="")
text(8,2.5,labels="eta = 25")
lines(CP, logfun(CP, 1,-10,.4),type="l", col="red")
text(8,2,labels="eta = 35")
lines(CP, logfun(CP, 1,-10,.3),type="l", col="blue")
text(8,1.5,labels="eta = 45")
lines(CP, logfun(CP, 1,-10,.2),type="l", col="purple")
text(8,1,labels="eta = 55")

```

#### Gamma responsivity to GDH given fixed parameters and varying CP rates
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 5}

p1 <- .00025
p2 <- -10
p3 <- .3
tit <- paste("alpha:", p1, "beta:", p2, "eta:", p3)
plot(GDH, logit_mak_gamma(GDH,CP = 50, p1,p2,p3), type="l", main=tit, ylab="")
text(18000,5,labels="CP = 50")
lines(GDH, logit_mak_gamma(GDH,CP = 45, p1,p2,p3), type="l",  col="red")
text(18000,4,labels="CP = 40")
lines(GDH, logit_mak_gamma(GDH,CP = 35, p1,p2,p3), type="l",  col="blue")
text(18000,3,labels="CP = 30")
lines(GDH, logit_mak_gamma(GDH,CP = 25, p1,p2,p3), type="l",  col="purple")
text(18000,.5,labels="CP = 20")
lines(GDH, logit_mak_gamma(GDH,CP = 15, p1,p2,p3), type="l",  col="grey")
text(18000,0,labels="CP = 15")

```

#### Simulated budbreak responsivity (proportion in [0,1]) to GDH given fixed parameters, varying CP rates, and arbitrary cutpoints (2,5)
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 5}

p1 <- .00025
p2 <- -10
p3 <- .3
tit <- paste("alpha:", p1, "beta:", p2, "eta:", p3)
plot(GDH, logit_mak(GDH,CP = 50, p1,p2,p3), type="l", main=tit, ylab="")
text(18000,.9,labels="CP = 50")
lines(GDH, logit_mak(GDH,CP = 45, p1,p2,p3), type="l",  col="red")
text(18000,.8,labels="CP = 40")
lines(GDH, logit_mak(GDH,CP = 35, p1,p2,p3), type="l",  col="blue")
text(18000,.5,labels="CP = 30")
lines(GDH, logit_mak(GDH,CP = 25, p1,p2,p3), type="l",  col="purple")
text(18000,.2,labels="CP = 20")
lines(GDH, logit_mak(GDH,CP = 15, p1,p2,p3), type="l",  col="grey")
text(18000,.1,labels="CP = 15")

```



# Model MCMC diagnostics: rhat (convergence) and effective sample ratio (sampling efficiency)
```{r}

#d <- read.csv("../datafiles/Raw_combined_excel.csv")
#varlist <- unique(d$variety)
#varlist <- varlist[varlist != "Kordia (M)"]

#for(i in 1:9) {
#load(paste0("../modelfits/",varlist[i],"_3p_updated"))
#rhat(fit.test) %>% plot(title=paste(varlist[i], "rhat"))
#neff_ratio(fit.test) %>% plot(title=paste(varlist[i], "neff"))
#}
```

All model fits can additionally be inspected with shinystan via launch_shinystan()


