---
title: "Model details"
author: "Graeme Baird"
date: "5/24/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../codefiles/Functions.R")
library(cowplot)
library(bayesplot)
library(magrittr)
```

### Underlying model details

In this paper we present an extension of a cumulative link ordinal regression model to include nonlinear effects, and it's application to determining chill sufficiency in cherry production. Below are figures illustrating the effect of variations in parameters. See the associated paper and model_math file for underlying calculations.


### Plotting response of nonlinear CP equation to changes in parameters (X:CP, Y:Coefficient output)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
out <- 100

CP <- seq(0,70,length.out=out)

```

Parameter variance within the non-linear component is best visualized by manipulation of individual parameters, as follows. Parameter ranges demonstrated below are appropriate for un-scaled variables and don't necessarily match actual model parameter estimates, which are fit on scaled observations. 


#### Changes in alpha determine asymptotic maximum possible growth rate 

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 5}

CP <- seq(0,70,length.out=100)
GDH <- seq(0,20000,length.out=out)

plot(CP, logfun(CP, 3,-30,1),type="l", xlim=c(0,80), ylab="")
text(75,3,labels="alpha = 3")
lines(CP, logfun(CP, 2,-30,1),type="l", col="red")
text(75,2,labels="alpha = 2")
lines(CP, logfun(CP, 1,-30,1),type="l", col="blue")
text(75,1,labels="alpha = 1")
lines(CP, logfun(CP, .5,-30,1),type="l", col="purple")
text(75,.5,labels="alpha = .5")

```

#### Changes in beta determine overall location of curve
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 5}

plot(CP, logfun(CP, 3,-40,1),type="l", xlim=c(0,70), ylab="")
text(45,2.5,labels="beta = -40")
lines(CP, logfun(CP, 3,-30,1),type="l", col="red")
text(35,2,labels="beta = -30")
lines(CP, logfun(CP, 3,-20,1),type="l", col="blue")
text(25,1.5,labels="beta = -20")
lines(CP, logfun(CP, 3,-10,1),type="l", col="purple")
text(15,1,labels="beta = -10")

```

#### Changes in eta determine location and acceleration between inflection points
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 5}

plot(CP, logfun(CP, 1,-10,.5),type="l", xlim=c(0,70), ylab="")
text(15,.8,labels="eta = .5")
lines(CP, logfun(CP, 1,-10,.4),type="l", col="red")
text(25,.6,labels="eta = .4")
lines(CP, logfun(CP, 1,-10,.3),type="l", col="blue")
text(30,.5,labels="eta = .3")
lines(CP, logfun(CP, 1,-10,.2),type="l", col="purple")
text(50,.4,labels="eta = .2")

```

#### Gamma responsivity to GDH given fixed parameters and varying CP rates
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 5}

p1 <- .00025
p2 <- -10
p3 <- .3
tit <- paste("alpha:", p1, "beta:", p2, "eta:", p3)
plot(GDH, logit_mak_gamma(GDH,CP = 50, p1,p2,p3), type="l", main=tit, ylab="")
text(18000,5,labels="CP = 50")
lines(GDH, logit_mak_gamma(GDH,CP = 45, p1,p2,p3), type="l",  col="red")
text(18000,4,labels="CP = 40")
lines(GDH, logit_mak_gamma(GDH,CP = 35, p1,p2,p3), type="l",  col="blue")
text(18000,3,labels="CP = 30")
lines(GDH, logit_mak_gamma(GDH,CP = 25, p1,p2,p3), type="l",  col="purple")
text(18000,.5,labels="CP = 20")
lines(GDH, logit_mak_gamma(GDH,CP = 15, p1,p2,p3), type="l",  col="grey")
text(18000,0,labels="CP = 15")

```

#### Simulated budbreak responsivity (proportion in [0,1]) to GDH given fixed parameters, varying CP rates, and arbitrary cutpoints (2,5)
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width = 5}

p1 <- .00025
p2 <- -10
p3 <- .3
tit <- paste("alpha:", p1, "beta:", p2, "eta:", p3)
plot(GDH, logit_mak(GDH,CP = 50, p1,p2,p3), type="l", main=tit, ylab="")
text(18000,.9,labels="CP = 50")
lines(GDH, logit_mak(GDH,CP = 45, p1,p2,p3), type="l",  col="red")
text(18000,.8,labels="CP = 40")
lines(GDH, logit_mak(GDH,CP = 35, p1,p2,p3), type="l",  col="blue")
text(18000,.5,labels="CP = 30")
lines(GDH, logit_mak(GDH,CP = 25, p1,p2,p3), type="l",  col="purple")
text(18000,.2,labels="CP = 20")
lines(GDH, logit_mak(GDH,CP = 15, p1,p2,p3), type="l",  col="grey")
text(18000,.1,labels="CP = 15")

```



# Model MCMC diagnostics: rhat (convergence) and effective sample ratio (sampling efficiency)
```{r}

#d <- read.csv("../datafiles/Raw_combined_excel.csv")
#varlist <- unique(d$variety)
#varlist <- varlist[varlist != "Kordia (M)"]

#for(i in 1:9) {
#load(paste0("../modelfits/",varlist[i],"_3p_updated"))
#rhat(fit.test) %>% plot(title=paste(varlist[i], "rhat"))
#neff_ratio(fit.test) %>% plot(title=paste(varlist[i], "neff"))
#}
```

All model fits can additionally be inspected with shinystan via launch_shinystan()


